Packaging research artefacts with RO-Crate

### 2022-01-02 latex/ro-crate.tex
Some typos. LanguageTag is not used.

Notable from this triple-like formalization is that a RO-Crate $R$ is
fully represented as a tree at depth 2 helped by the use of
%PDAG%
$\mathbh{lri}$
%PDAG%
nodes. For instance the aggregation from the root entity
$\textit{hasPart}(\texttt{{\ldots}interviews/>})$ is at same level as the data
entity's property \textit{license}(\texttt{{\ldots}CC-BY-\noxml{\break}NC-4.0>}) and that
contextual entity's attribute $ \textit{name}(\text{{\ldots}Non Commercial
4.0''})$. As shown in Section~\ref{sec:jsonld}, the RO-Crate Metadata
File serialization is an equivalent shallow tree, although at depth 3
to cater for the JSON-LD preamble of \texttt{"@context"} and \texttt{"@graph"}.

@@

A formalised RO-Crate in $\mathcal{L}_{\mathit{rocrate}}$ can be mapped to
different serializations.
Assume a simplified\footnote{This simplification does not cover the extensive list of literal
datatypes built-in to RDF 1.1, only strings and decimal real numbers.
Likewise, language of literals are not included.} language $\mathcal{L}_{\textit{RDF}}$
based on the RDF abstract syntax \cite{rdfworkinggroup_2014}:
%
{\fontsize{9.8}{11.8}\selectfont{\begin{eqnarray*}
\mathcal{L}_{\textit{RDF}} & \equiv& \bigl\{ \textit{Triple}(s,p,o),
\textit{IRI}(i), \textit{BlankNode}(b), \textit{Literal}(s), \mathbb{IRI},
\mathbb{S}, \mathbb{R} \bigr\}
\\
\mathbb{D}_{\textit{RDF}} & \equiv& \mathbb{S}
\\
\forall i \ \textit{IRI}(i) & \Rightarrow& i \in\mathbb{IRI}
\\
\forall s \forall p \forall o \ \textit{Triple}(s,p,o) & \Rightarrow& \bigl(
\textit{IRI}(s) \lor \textit{BlankNode}(s) \bigr) \land
\\
& & \textit{IRI}(p) \land
\\
& & \bigl(\textit{IRI}(o) \lor \textit{BlankNode}(o) \lor \textit{Literal}(o)
\bigr)
\\
\textit{Literal}(v) & \models& \textit{Value}(v) \land \textit{Datatype}(v,t)
\land \textit{IRI}(t)
\\
\forall v \ \textit{Value}(v) & \Rightarrow& v \in\mathbb{S}
\\
\textit{LanguageTag}(v, l) & \equiv& Datatype (v,
\\
&& \texttt{<http://www.w3.org/1999/02/22-rdf-syntax-ns\#langString>} )
\end{eqnarray*}}}

Below follows a mapping from $\mathcal{L}_{\mathit{rocrate}}$ to $\mathcal
{L}_{\textit{RDF}}$ using Schema.org as vocabulary:
%
{\fontsize{9.6}{11.6}\selectfont{\begin{eqnarray*}
\textit{Property}(p) & \Rightarrow& \textit{type}(p, \texttt
{<http://www.w3.org/2000/01/rdf-schema\#Property>})
\\
\textit{Class}(c) & \Rightarrow& \textit{type}(c, \texttt {<http://www.w3.org/2000/01/rdf-schema
\#Class>})
\\
\textit{Dataset}(d) & \Rightarrow& \textit{type}(d, \texttt {<http://schema.org/Dataset>})
\\
\textit{File}(f) & \Rightarrow& \textit{type}(f, \texttt {<http://schema.org/MediaObject>})
\\
\textit{ContextualEntity}(e) & \Rightarrow& \textit{type}(f, \texttt
{<http://schema.org/Thing>})
\\
\textit{CreativeWork}(e) & \Rightarrow& \textit{ContextualEntity}(e) \land
\textit{type}(e, \texttt{<http://schema.org/CreativeWork>})
\\
\textit{hasPart}(e, t) & \Rightarrow& \textit{Relation}(e, \texttt
{<http://schema.org/hasPart>}, t)
\\
\textit{name}(e, n) & \Rightarrow& \textit{Attribute}(e, \texttt
{<http://schema.org/name>}, n)
\\
\textit{description}(e, s) & \Rightarrow& \textit{Attribute}(e, \texttt
{<http://schema.org/description>}, s)
\\
\textit{datePublished}(e, d) & \Rightarrow& Attribute(e, \texttt
{<http://schema.org/datePublished>}, d)
\\
\textit{license}(e, l) & \Rightarrow& \textit{Relation}(e, \texttt
{<http://schema.org/license>}, l) \land \textit{CreativeWork}(l)
\\
\textit{type}(e, t) & \Rightarrow& \textit{Relation}(e, \texttt
{<http://www.w3.org/1999/02/22-rdf-syntax-ns\#type>}, t)
\\
& & {}\land \textit{Class}(t)
\\
\textit{String}(s) & \equiv& \textit{Value}(s) \land s \in\mathbb{S}
\\
\textit{String}(s) & \Rightarrow& \textit{Datatype}(s, \texttt
{<http://www.w3.org/2001/XMLSchema\#string>})
\\
\textit{Decimal}(d) & \equiv& \textit{Value}(d) \land d \in\mathbb{R}
\\
\textit{Decimal}(d) & \Rightarrow& \textit{Datatype}(d, \texttt
{<http://www.w3.org/2001/XMLSchema\#decimal>})
\\
\textit{Relation}(s,p,o) & \Rightarrow& \textit{Triple}(s,p,o) \land
\textit{IRI}(s) \land \textit{IRI}(o)
\\
\textit{Attribute}(s,p,o) & \Rightarrow& \textit{Triple}(s,p,o) \land
\textit{IRI}(s) \land \textit{Literal}(o)
\end{eqnarray*}}}

>>>>

Notable from this triple-like formalization is that a RO-Crate $R$ is
fully represented as a tree at depth 2 helped by the use of
%PDAG%
$\mathbh{IRI}$
%PDAG%
nodes. For instance the aggregation from the root entity
$\textit{hasPart}(\texttt{{\ldots}interviews/>})$ is at same level as the data
entity's property \textit{license}(\texttt{{\ldots}CC-BY-\noxml{\break}NC-4.0>}) and that
contextual entity's attribute $ \textit{name}(\text{{\ldots}Non Commercial
4.0''})$. As shown in Section~\ref{sec:jsonld}, the RO-Crate Metadata
File serialization is an equivalent shallow tree, although at depth 3
to cater for the JSON-LD preamble of \texttt{"@context"} and \texttt{"@graph"}.

@@

A formalised RO-Crate in $\mathcal{L}_{\mathit{rocrate}}$ can be mapped to
different serializations.
Assume a simplified\footnote{This simplification and mapping does not cover the extensive list of literal
datatypes built into RDF 1.1, only strings and decimal real numbers.
Likewise, \textit{LanguageTag} is deliberately not utillised below.} language $\mathcal{L}_{\textit{RDF}}$
based on the RDF abstract syntax \cite{rdfworkinggroup_2014}:
%
{\fontsize{9.8}{11.8}\selectfont{\begin{eqnarray*}
\mathcal{L}_{\textit{RDF}} & \equiv& \bigl\{ \textit{Triple}(s,p,o),
\textit{IRI}(i), \textit{BlankNode}(b), \textit{Literal}(s), \mathbb{IRI},
\mathbb{S}, \mathbb{R} \bigr\}
\\
\mathbb{D}_{\textit{RDF}} & \equiv& \mathbb{S}
\\
\forall i \ \textit{IRI}(i) & \Rightarrow& i \in\mathbb{IRI}
\\
\forall s \forall p \forall o \ \textit{Triple}(s,p,o) & \Rightarrow& \bigl(
\textit{IRI}(s) \lor \textit{BlankNode}(s) \bigr) \land
\\
& & \textit{IRI}(p) \land
\\
& & \bigl(\textit{IRI}(o) \lor \textit{BlankNode}(o) \lor \textit{Literal}(o)
\bigr)
\\
\textit{Literal}(v) & \models& \textit{Value}(v) \land \textit{Datatype}(v,t)
\land \textit{IRI}(t)
\\
\forall v \ \textit{Value}(v) & \Rightarrow& v \in\mathbb{S}
\\
\textit{LanguageTag}(v, l) & \equiv& Datatype (v,
\\
&& \texttt{<http://www.w3.org/1999/02/22-rdf-syntax-ns\#langString>} )
\end{eqnarray*}}}

Below follows a mapping from $\mathcal{L}_{\mathit{rocrate}}$ to $\mathcal
{L}_{\textit{RDF}}$ using Schema.org as vocabulary:
%
{\fontsize{9.6}{11.6}\selectfont{\begin{eqnarray*}
\textit{Property}(p) & \Rightarrow& \textit{type}(p, \texttt
{<http://www.w3.org/2000/01/rdf-schema\#Property>})
\\
\textit{Class}(c) & \Rightarrow& \textit{type}(c, \texttt {<http://www.w3.org/2000/01/rdf-schema\#Class>})
\\
\textit{Dataset}(d) & \Rightarrow& \textit{type}(d, \texttt {<http://schema.org/Dataset>})
\\
\textit{File}(f) & \Rightarrow& \textit{type}(f, \texttt {<http://schema.org/MediaObject>})
\\
\textit{ContextualEntity}(e) & \Rightarrow& \textit{type}(f, \texttt
{<http://schema.org/Thing>})
\\
\textit{CreativeWork}(e) & \Rightarrow& \textit{ContextualEntity}(e) \land
\textit{type}(e, \texttt{<http://schema.org/CreativeWork>})
\\
\textit{hasPart}(e, t) & \Rightarrow& \textit{Relation}(e, \texttt
{<http://schema.org/hasPart>}, t)
\\
\textit{name}(e, n) & \Rightarrow& \textit{Attribute}(e, \texttt
{<http://schema.org/name>}, n)
\\
\textit{description}(e, s) & \Rightarrow& \textit{Attribute}(e, \texttt
{<http://schema.org/description>}, s)
\\
\textit{datePublished}(e, d) & \Rightarrow& Attribute(e, \texttt
{<http://schema.org/datePublished>}, d)
\\
\textit{license}(e, l) & \Rightarrow& \textit{Relation}(e, \texttt
{<http://schema.org/license>}, l) \land \textit{CreativeWork}(l)
\\
\textit{type}(e, t) & \Rightarrow& \textit{Relation}(e, \texttt
{<http://www.w3.org/1999/02/22-rdf-syntax-ns\#type>}, t)
\\
& & {}\land \textit{Class}(t)
\\
\textit{String}(s) & \equiv& \textit{Value}(s) \land s \in\mathbb{S}
\\
\textit{String}(s) & \Rightarrow& \textit{Datatype}(s, \texttt
{<http://www.w3.org/2001/XMLSchema\#string>})
\\
\textit{Decimal}(d) & \equiv& \textit{Value}(d) \land d \in\mathbb{R}
\\
\textit{Decimal}(d) & \Rightarrow& \textit{Datatype}(d, \texttt
{<http://www.w3.org/2001/XMLSchema\#decimal>})
\\
\textit{Relation}(s,p,o) & \Rightarrow& \textit{Triple}(s,p,o) \land
\textit{IRI}(s) \land \textit{IRI}(o)
\\
\textit{Attribute}(s,p,o) & \Rightarrow& \textit{Triple}(s,p,o) \land
\textit{IRI}(s) \land \textit{Literal}(o)
\end{eqnarray*}}}
